Covariant Phase Space via MultiSymplectic Geometry

### 2017-06-16 cpsviams.tex
Little comment on vector bundles over Conf space.

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the sections of a suitable vector bundle composing with the jet functor $j$.
%
\begin{align*}
	D_\Lagrangian[\; ] &= 
	D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[ \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right) \\
	D_\Hamiltonian[\; ] &= 
	D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[ (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{align*}
%
The target base is a vector space, more precisely a Frechet space \cite[Cap.2]{Ban} and it makes sense to ask whether $D_\cdot [\phi]$ is equal to the 0-section.

@@

\section{Jacobi Operator}\label{Sec:JacobiOperator}
Intuitively, the Jacobi equations are the formal linearization of the equation of motion around a solution $\phi$.\\
Everything can be encoded in the following operator:
%
	\begin{definition}[Jacobi operator pertaining to $D$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp} is the bundle map over $F$:
		\begin{displaymath}
			J(\phi) \; : \; \phi^\ast(VF) \rightarrow	\phi^\ast(V^\TwistedLinear F) 
			\quad s.t. \quad  \left(J(\phi)\right) V
			= \sigma \left( \left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda](x) \right\rvert_{\lambda=0} \right)
		\end{displaymath}
		where $V\in (\phi^\ast(VF))_{\phi(x)} = V_{\phi(x)}F$ and $\phi_\lambda$ is a regular variation generated by a $\delta\phi$ such that $\delta\phi_x = V$ and
		$\sigma$ is the canonical bundle-morphism \footnote{see \ref{Section:sigma} for further details.}
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
	\end{definition}
	%
	\begin{remark}
		Being $J$ a bundle morphism, it can be naturally interpreted functionally as
		\begin{displaymath}
			\left(J(\phi)\right) [\: ] : 
			\Gamma^\infty\left(\phi^\ast(VF)\right) \rightarrow	
			\Gamma^\infty \left( 
			\phi^\ast(V^\TwistedLinear F)
			\right)
		\end{displaymath}
	\end{remark}
%
  \begin{minipage}[r]{0.5\textwidth}
    \includegraphics[width=0.85\textwidth]{Pictures/Jacobi.png}
  \end{minipage}\hfill
  \begin{minipage}[l]{0.5\textwidth}
	Note that $ \left(J(\phi)\right) [ \delta\phi ]$ and $D$ mapt to the same space.
  \end{minipage}

Since $D[\phi_\lambda](x)$ is a smooth curve on $ V^\TwistedLinear F $, that is also vertical w.r.t. $M$, 
the derivative involved has to be intended as the tangent velocity vector,
thus $\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]$ is the vertical vector field associated to the infinitesimal variation.\\
Exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$.
		Hence the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.

@@

\begin{lemma}\label{Prop:JacobieLie}
	The Jacobi operator $\left(J(\phi) \right) [\delta \phi]$
	around a solution $\phi$ when acting on an infinitesimal variation $\delta \phi \in T_\phi \Conf$
	via a twisted form that can be evaluated on any $V \in \Gamma\left(V F \right)$
	yields the following expression:
	\begin{displaymath}	
		\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle =
		\phi^\ast \; \Lie_{\delta \phi} \left( i_V \omega \right)
	\end{displaymath}
\end{lemma}
\begin{proof}
The explicit action of the Jacobi Operator can be expressed as follows:
\begin{displaymath}
	D[\phi_\cdot](\cdot): (\lambda, x^\mu) \mapsto
	\left( x^\mu, \phi_\lambda^A (x), \hat{D}[\phi_\lambda]_A(x) \right)
	\in V^\TwistedLinear F 
\end{displaymath}
%
\begin{align*}
	\frac{\partial}{\partial \lambda} \eval{D[\phi_\lambda](\cdot)}_{\lambda=0}:
	x \mapsto
	\biggr(& x^\mu, \phi_0^A(x), \hat{D}[\phi_0]_A(x),\\
	&	
	\eval{\dfrac{\partial \phi^A_\lambda(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q^A} + 
	\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q_A}	  \biggr)
	 \in V_M(V^\TwistedLinear F)
\end{align*}
\begin{align*}
%	\left(J(\phi)\right) [\: ] (\cdot) =
	\sigma \frac{\partial}{\partial \lambda} \eval{D[\phi_\lambda](\cdot)}_{\lambda=0}:
	x \mapsto &
	\left( x^\mu, \phi_0^A(x), 0,
	\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q_A}	  \right)
	 \in V_F(V^\TwistedLinear F) \\
	 & \qquad \qquad \qquad \vert\wr \\
	 &
		\left( x^\mu, \phi_0^A(x),
			\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0} \right)
	 		\in V^\TwistedLinear F 
\end{align*}
	Therefore for all $x\in M$:
	\begin{align*}
		\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle_x &=
		\left\langle \left(J(\phi) \right) [\delta \phi] (x) \right\vert
		\left. V_x \right\rangle =
		\left\langle\left.		
		\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
		\right\vert
		 V_x \right\rangle = \\
		 &=
		 \eval{
		 	\dfrac{\partial}{\partial \lambda}
		\left\langle\left.			 	
		 	\hat{D}[\phi_\lambda]
		 			\right\vert
		 V \right\rangle _x
		 }_{\lambda=0}
	\end{align*}
	from the linearity of the twisted duality pairing $<\,|\,>$.\\
	Plugging  the definition of motion operator $D$ in the latter expression, we obtain:
	\begin{displaymath}
				\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle =
		 \eval{
		 	\dfrac{\partial}{\partial \lambda}
				\left(
					\phi^\ast_\lambda \left( i_V \omega \right) \right)
		 }_{\lambda=0}	=
		 \phi_0^\ast
		 \eval{
				\dfrac{\partial}{\partial \lambda}
				\left(
			 		\left(\phi_\lambda \cdot \phi_0^{-1}\right)^\ast
					\left( i_V \omega \right) \right)
		 }_{\lambda=0}	=
		 		\phi_0^\ast \; \Lie_{\delta \phi} \left( i_V \omega \right)
	\end{displaymath}
	recognizing that $\left(\phi_\lambda \cdot \phi_0^{-1}\right)$ is the flow generated by $\delta \phi$.
\end{proof}

@@

We have the following short exact sequence of bundle morphisms over $W$:
	\begin{center}
	  \includestandalone[mode=buildnew]{Pictures/Figure_sigma_exact}
	\end{center}
inasmuch as
%
\begin{eqnarray*}
	\ker(i) = 0 (E) = i' \left( 0 \right)
	\ker\left( T \pi_{E,W} \right) = V_E W = i\left( V_E W \right)
	\ker\left(T \pi_{M,E}\right) = V_M E = T \pi_{E,W} \left( V_M W \right)
\end{eqnarray*}

We can exhibit a right splitting, that is:
\begin{displaymath}
	u : C \rightarrow B \quad \textrm{s.t.} \quad u \circ T \pi_{E,W} = \textrm{id}_B
\end{displaymath}
so, by splitting lemma, we have a left splitting $\sigma : B \rightarrow A$ that is exactly the desired operator.

The claimed right splitting is the following:
	\begin{displaymath}
		u = T0 \circ m \; : \; 
		\begin{array}{ccc}
			\pi^\ast_{E,W} \left(TE \right) &\rightarrow TE &\rightarrow TW \\
			\{w\} \times \{V_{\pi(w)} \} &\mapsto V_{\pi(w)} &\mapsto \left( T_{\pi(w)}0\right) (V_{\pi(w)})		
		\end{array}
	\end{displaymath}
%
In fact, since $0 \in \Gamma (W,E)$, we have:
\begin{displaymath}
	\textrm{T}_e 0 \circ \textrm{T}_\omega \pi_{E,W} V =
	\textrm{T}_\omega \left( 0 \circ \pi_{E,W} \right) V = 
	V \qquad \forall V \in \left( V_M W \right)_\omega ,\; \omega \in \pi^-1(e)
\end{displaymath}
Therefore, we have the following decomposition:

\begin{displaymath}
	\left( V_M W \right)_{0(e)} = 
	i\left[ V_E W \right]_{0(e)} \oplus \left( T_e 0 \left[ V_e E\right] \right)
\end{displaymath}
which determines an intrinsic mapping 
\begin{displaymath}
	\sigma_e : \left( V_M W \right)_{0(e)} \rightarrow \left( V_E W \right)_{0(e)} \simeq W_e
\end{displaymath}
that is simply the projection on the first component.

@@

thus proving the closure of the symplectic current.

>>>>

The key point of the latter definition is that the \emph{Euler-Lagrange} map and the \emph{DeDonder-Weyl} map, being bundle-morphism, can be regarded as mapping from $\Conf$ to the sections of a suitable vector bundle composing with the jet functor $j$.
%
\begin{align*}
	D_\Lagrangian[\; ] &= 
	D_\Lagrangian \circ j^2 : \quad \Gamma^\infty(E) \ni \varphi \quad \mapsto \quad D_\Lagrangian[ \varphi ] \in \Gamma^\infty \left( \phi^\ast \left(V^\TwistedLinear E \right)\right) \\
	D_\Hamiltonian[\; ] &= 
	D_\Hamiltonian \circ j^1 : \quad \Gamma^\infty(F) \ni (\varphi,\pi) \quad \mapsto \quad D_\Hamiltonian[ (\varphi,\pi) ] \in \Gamma^\infty \left( (\varphi,\pi)^\ast \left(V^\TwistedLinear F \right)\right)
\end{align*}
%
The target base is a vector space, more precisely a Frechet space \cite[Cap.2]{Ban}, and it makes sense to ask whether $D_\cdot [\phi]$ is equal to the 0-section.

@@

\section{Jacobi Operator}\label{Sec:JacobiOperator}
Intuitively, the Jacobi equations are the formal linearization of the equation of motion around a solution $\phi$.\\
Everything can be encoded in the following operator:
%
	\begin{definition}[Jacobi operator pertaining to $D$ around a solution $\phi \in \Sol$]\label{Def:JacobiOp} is the bundle map over $F$:
		\begin{displaymath}
			J(\phi) \; : \; \phi^\ast(VF) \rightarrow	\phi^\ast(V^\TwistedLinear F) 
			\quad s.t. \quad  \left(J(\phi)\right) V
			= \sigma \left( \left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda](x) \right\rvert_{\lambda=0} \right)
		\end{displaymath}
		where $V\in (\phi^\ast(VF))_{\phi(x)} = V_{\phi(x)}F$ and $\phi_\lambda$ is a regular variation generated by a $\delta\phi$ such that $\delta\phi_x = V$ and
		$\sigma$ is the canonical bundle-morphism \footnote{see \ref{Section:sigma} for further details.}
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
	\end{definition}
	%
	\begin{remark}
		Being $J$ a bundle morphism, it can be naturally interpreted functionally as
		\begin{displaymath}
			\left(J(\phi)\right) [\: ] : 
			\Gamma^\infty\left(\phi^\ast(VF)\right) \rightarrow	
			\Gamma^\infty \left( 
			\phi^\ast(V^\TwistedLinear F)
			\right)
		\end{displaymath}
	\end{remark}
%
	Note that $D[ \phi]$  and $ \left(J(\phi)\right) [ \delta \phi ]$  $(\forall \delta\phi \in T_\phi\Conf)$ are vectors in the same vector space.
	%
	Considering the "formal" infinite dimensional vector bundle 
	\begin{displaymath}
		\mathcal{E} =\bigsqcup_{\phi \in \Conf} 
			\Gamma^\infty \left( 
				\phi^\ast(V^\TwistedLinear F)
			\right)
			\rightarrow \Conf
	\end{displaymath}
	$D$ could be regarded as a section on it and $J$ as a bundle map from $T\Conf$ to $V\mathcal{E}\simeq \mathcal{E}$.
	\begin{center}
	  	\includestandalone[width=0.85\textwidth]{Pictures/Figure_Jacobi}
	\end{center}

Since $D[\phi_\lambda](x)$ is a smooth curve on $ V^\TwistedLinear F $, that is also vertical w.r.t. $M$, 
the derivative involved has to be intended as the tangent velocity vector,
thus $\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]$ is the vertical vector field associated to the infinitesimal variation.\\
Exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\diff}{\diff \lambda} D [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$.\\
		Hence the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.

@@

\begin{lemma}\label{Prop:JacobieLie}
	The Jacobi operator $\left(J(\phi) \right) [\delta \phi]$
	around a solution $\phi$ when acting on an infinitesimal variation $\delta \phi \in T_\phi \Conf$
	gives a twisted form that can be evaluated on any $V \in \Gamma\left(V F \right)$
	yielding the following expression:
	\begin{displaymath}	
		\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle =
		\phi^\ast \; \Lie_{\delta \phi} \left( i_V \omega \right)
	\end{displaymath}
\end{lemma}
\begin{proof}
The explicit action of the Jacobi Operator can be expressed as follows:
\begin{displaymath}
	D[\phi_\cdot](\cdot): (\lambda, x^\mu) \mapsto
	\left( x^\mu, \phi_\lambda^A (x), \hat{D}[\phi_\lambda]_A(x) \right)
	\in V^\TwistedLinear F 
\end{displaymath}
%
\begin{align*}
	\frac{\partial}{\partial \lambda} \eval{D[\phi_\lambda](\cdot)}_{\lambda=0}:
	x \mapsto
	\biggr(& x^\mu, \phi_0^A(x), \hat{D}[\phi_0]_A(x),\\
	&	
	\eval{\dfrac{\partial \phi^A_\lambda(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q^A} + 
	\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q_A}	  \biggr)
	 \in V_M(V^\TwistedLinear F)
\end{align*}
\begin{align*}
%	\left(J(\phi)\right) [\: ] (\cdot) =
	\sigma \frac{\partial}{\partial \lambda} \eval{D[\phi_\lambda](\cdot)}_{\lambda=0}:
	x \mapsto &
	\left( x^\mu, \phi_0^A(x), 0,
	\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
	 \dfrac{\partial}{\partial q_A}	  \right)
	 \in V_F(V^\TwistedLinear F) \\
	 & \qquad \qquad \qquad \vert\wr \\
	 &
		\left( x^\mu, \phi_0^A(x),
			\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0} \right)
	 		\in V^\TwistedLinear F 
\end{align*}
	Therefore for all $x\in M$:
	\begin{align*}
		\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle_x &=
		\left\langle \left(J(\phi) \right) [\delta \phi] (x) \right\vert
		\left. V_x \right\rangle =
		\left\langle\left.		
		\eval{\dfrac{\partial \hat{D}[\phi_\lambda]_A(x)}{\partial \lambda}}_{\lambda=0}
		\right\vert
		 V_x \right\rangle = \\
		 &=
		 \eval{
		 	\dfrac{\partial}{\partial \lambda}
		\left\langle\left.			 	
		 	\hat{D}[\phi_\lambda]
		 			\right\vert
		 V \right\rangle _x
		 }_{\lambda=0}
	\end{align*}
	from the linearity of the twisted duality pairing $<\,|\,>$.\\
	Plugging  the definition of motion operator $D$ in the latter expression, we obtain:
	\begin{displaymath}
				\left\langle \left(J(\phi) \right) [\delta \phi] \right\vert
		\left. V \right\rangle =
		 \eval{
		 	\dfrac{\partial}{\partial \lambda}
				\left(
					\phi^\ast_\lambda \left( i_V \omega \right) \right)
		 }_{\lambda=0}	=
		 \phi_0^\ast
		 \eval{
				\dfrac{\partial}{\partial \lambda}
				\left(
			 		\left(\phi_\lambda \cdot \phi_0^{-1}\right)^\ast
					\left( i_V \omega \right) \right)
		 }_{\lambda=0}	=
		 		\phi_0^\ast \; \Lie_{\delta \phi} \left( i_V \omega \right)
	\end{displaymath}
	recognizing that $\left(\phi_\lambda \cdot \phi_0^{-1}\right)$ is the flow generated by $\delta \phi$.
\end{proof}

@@

We have the following short exact sequence of bundle morphisms over $W$:
	\begin{center}
	  \includestandalone[mode=buildnew]{Pictures/Figure_sigma_exact}
	\end{center}
inasmuch as
%
\begin{align*}
	&\ker(i) = 0 (E) = i' \left( 0 \right) \\
	&\ker\left( T \pi_{E,W} \right) = V_E W = i\left( V_E W \right) \\
	&\ker\left(T \pi_{M,E}\right) = V_M E = T \pi_{E,W} \left( V_M W \right) \\
\end{align*}

%
We can exhibit a right splitting, that is:
\begin{displaymath}
	u : C \rightarrow B \quad \textrm{s.t.} \quad u \circ T \pi_{E,W} = \textrm{id}_B
\end{displaymath}
so, by splitting lemma, we have a left splitting $\sigma : B \rightarrow A$ that is exactly the desired operator.

The claimed right splitting is the following:
	\begin{displaymath}
		u = T0 \circ m \; : \; 
		\begin{array}{ccc}
			\pi^\ast_{E,W} \left(TE \right) &\rightarrow TE &\rightarrow TW \\
			\{w\} \times \{V_{\pi(w)} \} &\mapsto V_{\pi(w)} &\mapsto \left( T_{\pi(w)}0\right) (V_{\pi(w)})		
		\end{array}
	\end{displaymath}
%
In fact, since $0 \in \Gamma (W,E)$, we have:
\begin{displaymath}
	\textrm{T}_e 0 \circ \textrm{T}_\omega \pi_{E,W} V =
	\textrm{T}_\omega \left( 0 \circ \pi_{E,W} \right) V = 
	V \qquad \forall V \in \left( V_M W \right)_\omega ,\; \omega \in \pi^-1(e)
\end{displaymath}
Therefore, we have the following decomposition:
%
\begin{displaymath}
	\left( V_M W \right)_{0(e)} = 
	i\left[ V_E W \right]_{0(e)} \oplus \left( T_e 0 \left[ V_e E\right] \right)
\end{displaymath}
which determines an intrinsic mapping 
\begin{displaymath}
	\sigma_e : \left( V_M W \right)_{0(e)} \rightarrow \left( V_E W \right)_{0(e)} \simeq W_e
\end{displaymath}
that is simply the projection on the first component.

@@

thus the closure of the symplectic current is verified.
